---
title: "p8105_hw2_jy3269"
author: "Jingyi Yao"
date: "`r Sys.Date()`"
output: github_document
---

```{r,warning=FALSE,message=FALSE}
library(tidyverse)
```


## Problem 1

### 1. import csv data
```{r, warning=TRUE,message=FALSE}
transit <- read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() 

transit

```

### 2. clean data
```{r}
transit <- transit %>%  
  select(line, station_name, station_latitude,station_longitude, route1:route11, entry, vending, entrance_type, ada) %>% 
  mutate(entry=recode(entry,"YES"= TRUE,"NO"= FALSE))

transit

```

*  Description of the data:
  * the transit data set has : `r ncol(transit)` columns(variables) and `r nrow(transit)` rows
  * the dimenison is : `r dimension(transit)`
  * the variables are : `r names(transit)`


*  data cleaning steps:
  * unify variable name format using `janitor::clean_names`
  * select the required columns using `select()` in `dplyr`
  * convert the character values in `entry` column into logical values (TRUE or FALSE) using `mutate()` and `recode()`
  
  


```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```